Master thesis problem statement: Keyword search over heterogeneous sources

### 2012-10-09 doc/thesis/latex/literature_review.tex
some changes to report (still early)

\documentclass[a4paper,11pt,draft]{article}
\special{papersize=210mm,297mm}

% twocolumn
%\usepackage{fullpage} 
\usepackage[cm]{fullpage}
%\usepackage[margin=0.5in]{geometry}
%\usepackage{times}
\usepackage{color}
%\usepackage[small,compact]{titlesec}

\bibliographystyle{alpha}

% usually first paragraph is NOT indented, so this is commented out: 
% \usepackage{indentfirst}

\begin{document}

% Article top matter
\title{Master thesis problem statement: Keyword search over heterogeneous sources} %\LaTeX is a macro for printing the Latex logo
\author{Vidmantas Zemleris}  %\texttt formats the text to a typewriter style font
\date{\today}  %\today is replaced with the current date

%\maketitle

%
 \centerline{\Large \bf Searching heterogeneous data-sources: Master thesis problem statement} %% Paper title
 \medskip

 \centerline{Vidmantas Zemleris, \today}  %% Author name(s)
 \medskip

\section{Introduction}
At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
\footnote{For instance, at CERN,  due to many reasons (e.g. research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}. Further high turnover makes it harder extending these systems}. As data stored on one system may be related to data residing on other systems%
	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.

Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema. A data integration system based on simple structured queries is already in place. Still {\color{red}various improvements including support} for less restricted keyword queries and improvements to system's usability and performance still have to be researched.

% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating full set of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

For more use-cases of data retrieval at CMS Experiment see \cite{CMS_data08}.

%Following a similar approach a 
\subsection*{The Data Aggregation System}

The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent uses.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators. The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:

{\footnotesize 
\begin{verbatim}
dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
\end{verbatim}
}

The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.

Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g. program doing physics analysis or automatic software release validation).

% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}

\section{Problem statement}

\textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}

\subsection{Ease of use: Query Language over the mediated schema}
	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources). 
   So some sort of guiding helping user to build the query shall be useful.

	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.

Also to consider:
\begin{itemize}         
       \item input ambiguity, typos
       \item {\color{red} what about ranked search?}
\end{itemize}

\subsection{Performance}
\subsubsection{Handling distributed search efficiently}

\subsubsection{Other performance issues}
	     \begin{itemize}
	     					\item  {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)

	                		\item more intelligent caching needed:
	                			\begin{itemize}
	                			\item given an entity received from provider, determining if it is useful to cache for long-term
	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
	                			\item could make use of even older information warning user
	                			\item could also have different validity dates for certain fields. if certain field is not explicitly requested, even a very old cache could be used.
	                			\end{itemize}

	                		\item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
	                		\item pre-fetching common (sub-)queries: determining manually and/or automatically
	                		\item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.
                        \item ? with current API's for aggregation service has to fetch ALL records not just the first page..
                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item ? showing partial results
        \end{itemize}

@@

\subsection{Data integration}
\subsubsection*{Virtual integration}
In Virtual Integration, multiple data-sources are integrated with help of a mediator containing source descriptions and probably some schema mappings, while data is left at the sources. This is mainly the approach taken by our DAS.

\subsubsection*{Large scale search}
[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google) Virtual integration vs. surfacing. They also present ways for integrating systems without human intervention through use of statistical 'mediator'.

@@

{\color{red} In terms of virtual integration of services via system like our DAS, one could employ query cleaning process to suggest possible queries.

Inverted index could be built with some powerful full-text search engine (e.g. Xapian) and used to generate mapping to structured queries}

@@

\subsection{Searching 'Deep Web' and Heterogeneous web services}
{\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}

In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.

The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}

%{\color{red}Check deep web (also recent research exist)}
%{\color{red}Performance: Answering Queries using Views}

\begin{verbatim}
Some other approaches could include:

- creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
(Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)

- (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007)
 : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases
\end{verbatim}

\section{Possible approaches}

\subsection{Ease of use}
\begin{verbatim}
- javascript interpreter
    : field + description -- dropdown searcheable list

- if entered keyword query suggest a structured query
	based on 'Query forms' and 'Keyword cleaning' approaches
	--> evaluate different approaches (HMM, etc) and implementations
	optionally with description in natural language
- select entity (e.g. dataset) --> give all available keys from forms, however there are still combinations which are enforced
\end{verbatim}
%(could use Natural Language Generation)

\subsection{Performance}
\subsubsection{Continuous view maintenance at Data providers with large DBs}
Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

\subsection{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs

%\newline
integration at DAS level. (at source DB could be more performant a little)

\subsection{More intelligent caching}
Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).

Cached copy may be shown before hand while up to date results could be retrieved on user's request.

\textbf{Decide if to cache or not...}

An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.

\begin{verbatim}
Also: Caching of intermediary results, not only the queries
\end{verbatim}

\section{Work status}
TODO

\begin{verbatim}
- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
- preliminary literature review (to be continued deeper + waiting for book arrival: Principles of Data Integration)

--- some analysis of DAS logs

Upcomming Work items:
- couple of fairly simple prototypes of UI/access patters for simpler DAS querying
- check on performance improvements after creation of materialized view(s)
- interview (more) DAS users
\end{verbatim}

% ---  Bibliography
\bibliographystyle{unsrt}
\addcontentsline{toc}{chapter}{Bibliography}

%\nocite{*}
\thispagestyle{empty}
\bibliography{refs}

\pagebreak
\appendix
\section{DAS Query logs}

non parsible queries: 98923

\subsection*{Most common query patterns (Feb 2012-June 2012)}
\begin{verbatim}
total valid queries: 569408

50.16% (285605) :	dataset dataset.name=?
13.54% (77071)  :	site dataset.name=?
8.96%  (51035) 	:	file dataset.name=?
5.88%  (33504) 	:	run dataset.name=?
2.59%  (14739) 	:	release dataset.name=?
2.11%  (12036) 	:	config dataset.name=?
1.65%  (9420) 	:	dataset run.run_number=?
1.34%  (7642) 	:	block dataset.name=?
1.24%  (7084) 	:	dataset site.name=?
1.15%  (6562) 	:	dataset dataset.name=? release.name=?
1.10%  (6287) 	:	parent dataset.name=?
0.96%  (5488) 	:	file file.name=?
0.92%  (5245) 	:	dataset dataset.name=? status.name=?
0.77%  (4363) 	:	dataset release.name=?
0.75%  (4257) 	:	file dataset.name=? run.run_number=?
0.68%  (3874) 	:	run run.run_number=?
0.53%  (2994) 	:	site site.name=?
0.45%  (2576) 	:	site file.name=?
0.45%  (2556) 	:	dataset file.name=?
0.43%  (2438) 	:	lumi file.name=?
0.40%  (2282) 	:	dataset dataset.name=? site.name=?
0.35%  (1999) 	:	file block.name=?
0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
0.29%  (1640) 	:	group dataset.name=?
0.29%  (1631) 	:	lumi run.run_number=?
0.25%  (1402) 	:	child dataset.name=?
0.22%  (1268) 	:	run file.name=?
0.21%  (1204) 	:	block block.name=?
0.19%  (1088) 	:	release release.name=?
0.16%  (936) 	:	site block.name=?
0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
0.10%  (594) 	:	parent file.name=?
0.08%  (455) 	:	dataset block.name=?
0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
0.06%  (360) 	:	file dataset.name=? site.name=?

Interesting non structured queries:
py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
T!_CERN
*herwig*/AODSIM

\end{verbatim}

\end{document}

>>>>

@@

\subsection{Searching 'Deep Web' and Heterogeneous web services}

In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach. The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}

\subsubsection*{Deep web search at Google}
[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google) Virtual integration vs. surfacing. They also present ways for integrating systems without human intervention through use of statistical 'mediator'.

@@

@@

% {\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}

% Some other approaches could include: - creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
% (Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)

% TODO: (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007) : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases

### 2012-10-09 doc/thesis/latex/report_sept.tex
some changes to report (still early)

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating full set of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

@@

The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent uses.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators. The combined query results could be later 'pipped' for further filtering and aggregation (min, avg, etc), e.g.:

@@

\section{Problem statement}
\subsection{Ease of use: Query Language over the mediated schema}
	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources). 
   So some sort of guiding helping user to build the query shall be useful.

@@

Also to consider:
\begin{itemize}         
       \item input ambiguity, typos
       \item {\color{red} what about ranked search?}
\end{itemize}

\subsection{Performance}
\subsubsection{Handling distributed search efficiently}

\subsubsection{Other performance issues}
	     \begin{itemize}
	                		\item more intelligent caching needed:
	                			\begin{itemize}
	                			\item given an entity received from provider, determining if it is useful to cache for long-term
	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
	                			\item could make use of even older information warning user
	                			\item could also have different validity dates for certain fields. if certain field is not explicitly requested, even a very old cache could be used.
	                			\end{itemize}
	                		\item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.
	                		\item pre-fetching common (sub-)queries: determining manually and/or automatically
	                		\item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.
                        \item ? with current API's for aggregation service has to fetch ALL records not just the first page..
                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item ? showing partial results
        \end{itemize}

\section{Preliminary Literature review}

% Overview: keyword search is good for non-structured documents, it is not [as] effective with structured sources\cite{Levy96}, therefore the keyword search on relational databases approach is good then no specific APIs exist and it's there are no resources to do \textbf{a full integration} .

\textbf{\color{red}TODO: Overview}

\subsection{Keyword search over DBs}
The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade. There exist many papers and a couple of quite comprehensive PhD. dissertations\cite{PhD_2011, PhD_2012} describing performance optimization details, etc.

There was a number of variations of the problem including returning of only Top-k results expecting that users only care about the quality of limited number of results but not completeness of results, ... \textbf{\color{red}describe this in details}

\subsubsection*{Ranking Query Templates based on keyword query}
A simple way to access relational database could  be through a set of predefined named query templates (SQL with selection parameters or operators still to be specified) exposed to a user as a Form that the user has to fill in.

\cite{forms_kws} proposes alternative approach for processing keyword queries over relational databases: given a keyword query, instead of returning database tuples one could rank query forms that best matches the query for user to choose the right one (if they are properly named this is fairly easy). The ranking is done by \textbf{\color{red}TODO}.

An interesting feature of this approach is that a Query Template is functionally similar to any autonomous web service (which given the parameters would in turn execute that query on its database).  In case of the Data Aggregation System, a user after entering a keyword query could be provided with a ranked list of structured queries (attribute=value) that could be processed given data source constraints (e.g. parameters required) and if needed he could refine his search (e.g. provide more parameters).

\subsubsection*{Semantic approach}

Researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users in collaboration with \textit{Credit Suisse}.  To converting natural language queries (that in addition to keyword search could convey some semantic structure) into SQL statements they employed some natural language processing and used meta-data describing the schema and the domain (ontologies at multiple levels).

Even on a large data-warehouse of ~220GB data with a complex schema of 300 tables they reported fairly promising results that proves that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.

\subsubsection*{Automatic mapping between distributed services}
{\color{red}an interesting approach but not for CERN...}

\subsubsection*{\color{red}Question Answering}

% See Appendix #1, for evaluation of they Demo system.

\subsection{Searching Heterogeneous web services}
{\color{red}There doesn't seem to be much of recent papers going towards this direction (except from specific search systems like flight fare comparisons), however these are based on structured arguments.}

In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach.

The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}

{\color{red}Performance: Answering Queries using Views}

\begin{verbatim}
Some other approaches could include:

- creating virtual documents offline by joining tables, for instantaneous search results by employing IR techniques
(Indexing Relational Database Content Offline for Efficient Keyword-Based Search, 2003)

- (Efficient Keyword Search Across Heterogeneous Relational Databases, 2007)
 : combines schema matching and structure discovery techniques to ﬁnd approximate foreign-key joins across heterogeneous databases
\end{verbatim}

\section{Possible approaches}

\subsection{Ease of use}
\begin{verbatim}
- javascript interpreter
    : field + description -- dropdown searcheable list

- if entered keyword query suggest a structured query, optionally with description in natural language
- select entity (e.g. dataset) --> give all available keys from forms, however there are still combinations which are enforced
\end{verbatim}
%(could use Natural Language Generation)

\subsection{Performance}
\subsubsection{Continuous view maintenance at Data providers with large DBs}
Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

\subsection{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs

%\newline
integration at DAS level. (at source DB could be more performant a little)

\subsection{More intelligent caching}
Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).

Cached copy may be shown before hand while up to date results could be retrieved on user's request.

\textbf{Decide if to cache or not...}

An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.

\begin{verbatim}
Also: Caching of intermediary results, not only the queries
\end{verbatim}

@@

>>>>

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

@@

The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent use. DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators. The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:

@@

\section{Problem statement}

DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g. new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g. datasets stored at DBS do not change often). 
Further as the total data that is being queried could be fairly large (with growth of order of 1TB per year) and the data-services may have certain constraints (e.g. only certain queries allowed) as their main goal is supporting production (data taking from the CMS detector).

Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get up to date results.
It is also not less important to provide users with easy to use interface having a fairly flat learning curve, while at the same time not loosing the functionality of fairly complex queries. % expressive power.
\textbf{\color{red}Another issue is about handling distributed queries, where either selection criteria or selection values could come from distributed sources -- the biggest problem could be in the APIs they [DO NOT] provide}.

%TODO: Further it has to comply with dataset constraints

% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}

\subsection{Ease of use: Query Language over the mediated schema}
	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources). 
   So some sort of guiding helping user to build the query shall be useful.

   Supporting non-structured keyword queries is also worth investigation as many users reporting they are missing this Google-like search experience. {\color{red}Because DAS uses Boolean IR we could only rank specific structured queries.}

@@

Also to consider: input ambiguity and typos, some queries are more common than others.

\subsection{Performance}

\begin{itemize}
	     					\item  Query prioritization: {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)

                		\item Explore more intelligent caching

	                		% TODO: \item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.

					\item Could displaying Partial Results improve performance?
					\begin{itemize}
					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item {\color{red}e.g. intermediary aggregation results while still calculating}

					\end{itemize}

	                	 \item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.                        
\end{itemize}

\subsection{Handling distributed search efficiently}

\pagebreak
\include{proposed_solutions}

\pagebreak
\include{literature_review}

@@

\pagebreak
\appendix
\section{DAS Query logs}

non parsible queries: 98923

\subsection*{Most common query patterns (Feb 2012-June 2012)}
\begin{verbatim}
total valid queries: 569408

50.16% (285605) :	dataset dataset.name=?
13.54% (77071)  :	site dataset.name=?
8.96%  (51035) 	:	file dataset.name=?
5.88%  (33504) 	:	run dataset.name=?
2.59%  (14739) 	:	release dataset.name=?
2.11%  (12036) 	:	config dataset.name=?
1.65%  (9420) 	:	dataset run.run_number=?
1.34%  (7642) 	:	block dataset.name=?
1.24%  (7084) 	:	dataset site.name=?
1.15%  (6562) 	:	dataset dataset.name=? release.name=?
1.10%  (6287) 	:	parent dataset.name=?
0.96%  (5488) 	:	file file.name=?
0.92%  (5245) 	:	dataset dataset.name=? status.name=?
0.77%  (4363) 	:	dataset release.name=?
0.75%  (4257) 	:	file dataset.name=? run.run_number=?
0.68%  (3874) 	:	run run.run_number=?
0.53%  (2994) 	:	site site.name=?
0.45%  (2576) 	:	site file.name=?
0.45%  (2556) 	:	dataset file.name=?
0.43%  (2438) 	:	lumi file.name=?
0.40%  (2282) 	:	dataset dataset.name=? site.name=?
0.35%  (1999) 	:	file block.name=?
0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
0.29%  (1640) 	:	group dataset.name=?
0.29%  (1631) 	:	lumi run.run_number=?
0.25%  (1402) 	:	child dataset.name=?
0.22%  (1268) 	:	run file.name=?
0.21%  (1204) 	:	block block.name=?
0.19%  (1088) 	:	release release.name=?
0.16%  (936) 	:	site block.name=?
0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
0.10%  (594) 	:	parent file.name=?
0.08%  (455) 	:	dataset block.name=?
0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
0.06%  (360) 	:	file dataset.name=? site.name=?

Interesting non structured queries:
T!_CERN
*herwig*/AODSIM

\end{verbatim}
% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
